Building a continuous benchmarking ecosystem in bioinformatics

### 2025-04-23 0_original_submission/letter_to_editor.tex
before generating the bbl

\newcommand{\submission}{X}
\newcommand{\journal}{Y}
\newcommand{\submitter}{Z}

@@

\noindent Dear Editor,\\

@@

In this manuscript we describe the desired properties of benchmarking systems, e.g., software solutions to facilitate, standardize, coordinate and help distributing benchmarks.\\

We think our paper matches \journal's call for .\\

@@

\begin{itemize}
    \item Dr Salvador Capella Gutierrez, \texttt{salvador.capella@bsc.es}, leader of the Spanish National Bioinformatics (INB) Coordination Node at the Barcelona Supercomputing Center (BSC).
    \item Dr Malte Luecken, \texttt{malte.luecken@helmholtz-muenchen.de}, leader of the Computational Cellular Health Group, Helmholtz Munich.
    \item Prof Dr Carole Goble , \texttt{carole.goble@manchester.ac.uk}, full professor and co-leader of the Information Management Group, School of Computer Science in the University of Manchester.
    \item Prof Dr Johannes Koester, \texttt{johannes.koester@uni-due.de}, professor for Bioinformatics and Computational Oncology, Institute for AI in Medicine (IKIM), University Medicine Essen.
    \item Prof Dr Anne-Laure Boulesteix, \texttt{boulesteix@ibe.med.uni-muenchen.de}, leader of the Biometry in Molecular Medicine group, LMU Munich
\end{itemize}

>>>>

\newcommand{\submission}{Building a continuous benchmarking ecosystem in bioinformatics}
\newcommand{\journal}{NAR Genomics and Bioinformatics}
\newcommand{\submitter}{Izaskun Mallona and Mark D. Robinson}
\newcommand{\editor}{Dr Ana Rojas}

@@

Dear \editor,\\

@@

In this manuscript we argue benchmarking can benefit from benchmarking systems, e.g., software solutions to facilitate, standardize, coordinate and help distributing benchmarks, and discuss the desired properties of such benchmarking systems.\\

We think our paper matches \journal's scope and interest in robust and reproducible benchmarking of genomics and bioinformatics methods.\\

Our manuscript is available as a preprint at \url{https://arxiv.org/abs/2409.15472}.\\

@@

\begin{itemize}
    \item Dr Salvador Capella Gutierrez, \texttt{salvador.capella@bsc.es}, leader of the Spanish National Bioinformatics (INB) Coordination Node at the Barcelona Supercomputing Center (BSC).
    \item Dr Malte Luecken, \texttt{malte.luecken@helmholtz-muenchen.de}, leader of the Computational Cellular Health Group, Helmholtz Munich.
    \item Prof Dr Carole Goble , \texttt{carole.goble@manchester.ac.uk}, full professor and co-leader of the Information Management Group, School of Computer Science in the University of Manchester.
    \item Prof Dr Johannes Koester, \texttt{johannes.koester@uni-due.de}, professor for Bioinformatics and Computational Oncology, Institute for AI in Medicine (IKIM), University Medicine Essen.
    \item Prof Dr Anne-Laure Boulesteix, \texttt{boulesteix@ibe.med.uni-muenchen.de}, leader of the Biometry in Molecular Medicine group, LMU Munich.
    \item Prof Dr. Serguei Mangul, \texttt{serghei.mangul@gmail.com}, assistant professor of clinical pharmacy, University of Southern California.
    \item Prof Dr. Jean Yang, \texttt{jean.yang@sydney.edu.au}, professor, School of Mathematics and Statistics of the University of Sydney.
    \item Prof. Dr. Gregory Grann, \texttt{ggrant@pennmedicine.upenn.edu}, associate professor of genetics, Perelman School of Medicine
\end{itemize}

### 2025-04-23 1_first_review/manuscript.tex
before generating the bbl

% Emerald Publishing - Construction Innovation Submission Template
% by Oleksandr Melnyk
% Ver 0.0.4
% Based on: https://www.emeraldgrouppublishing.com/journal/ci#author-guidelines

@@

\begin{figure}
    \centering
    \includegraphics[width=\linewidth]{review_cycle_1/figures/fig_benchmarking_aligned.pdf}
     \caption[Run types, properties, capabilities and features.]{Solo and community benchmarks, workflows, and benchmarking-system-enabled (\textit{bs}) benchmarks all differ in run properties and their consequences in terms of capabilities and features, including whether: inputs, truths, hidden truths, metrics or run results are available for readers; code is free software, software environments are managed, and execution automated; there is a clear track of (human) contributors; results are run in comparable hardware; an entrypoint to results or result interpretation are provided. These run-intrinsic properties have consequences in terms of features (i.e., transparency, forkability, availability for metaanalysis) and technical capabilities (i.e. software versioning or ability to run locally without leaving traces - in \textit{incognito} mode). Green, blue and red ticks depict available, perhaps available and unavailable features, respectively. \label{fig:definitions}}

@@

\begin{figure}
    \centering
    \includegraphics[width=0.8\linewidth]{../original_submission/figures/bens_layers.pdf}
    \caption[Benchmarking conceptual layers and their challenges.]{Benchmarking conceptual layers and their challenges. The process of benchmarking is multilayered, with layer-specific factors belonging to different categories (colored boxes) and challenges (in red). \emph{Hardware}: infrastructure and its costs. \emph{Data}: dataset archival, openness, interoperability, and selection.  \emph{Software}: method implementations, reproducibility, workflow execution, continuous integration and delivery (CI/CD), versioning, quality assurance (QA). \emph{Community}: Standardization and impartiality (governance, transparency, building trust, long term maintainability). \emph{Knowledge}: research and meta-research, academic publications.}
    \label{fig:layers}
\end{figure}

@@

\begin{figure}
    \centering
        %\includegraphics[width=0.75\linewidth]{../original_submission/figures/bens_formalization.pdf}
    \includegraphics[width=0.75\linewidth]{figures/bens_formalization_arrows.pdf} % alternative
    \caption[Benchmark formalization.]{Benchmark formalization. \textbf{A}) Running a benchmark requires running a workflow (execution phase) to generate results to be critically evaluated (analysis phase). \textbf{B}) the execution phase consists on a mapping of methods to specific input files to generate output files, with optional steps to harmonize data and parametrize runs.}
    \label{fig:formalization}
\end{figure}

>>>>

1_% Emerald Publishing - Construction Innovation Submission Template
% by Oleksandr Melnyk
% Ver 0.0.4
% Based on: https://www.emeraldgrouppublishing.com/journal/ci#author-guidelines

@@

\begin{figure}
    \centering
    \includegraphics[width=\linewidth]{figures/fig1_fig_benchmarking_aligned.pdf}
     \caption[Run types, properties, capabilities and features.]{Solo and community benchmarks, workflows, and benchmarking-system-enabled (\textit{bs}) benchmarks all differ in run properties and their consequences in terms of capabilities and features, including whether: inputs, truths, hidden truths, metrics or run results are available for readers; code is free software, software environments are managed, and execution automated; there is a clear track of (human) contributors; results are run in comparable hardware; an entrypoint to results or result interpretation are provided. These run-intrinsic properties have consequences in terms of features (i.e., transparency, forkability, availability for metaanalysis) and technical capabilities (i.e. software versioning or ability to run locally without leaving traces - in \textit{incognito} mode). Green, blue and red ticks depict available, perhaps available and unavailable features, respectively. \label{fig:definitions}}

@@

\begin{figure}
    \centering
    \includegraphics[width=0.8\linewidth]{figures/fig2_bens_layers.pdf}
    \caption[Benchmarking conceptual layers and their challenges.]{Benchmarking conceptual layers and their challenges. The process of benchmarking is multilayered, with layer-specific factors belonging to different categories (colored boxes) and challenges (in red). \emph{Hardware}: infrastructure and its costs. \emph{Data}: dataset archival, openness, interoperability, and selection.  \emph{Software}: method implementations, reproducibility, workflow execution, continuous integration and delivery (CI/CD), versioning, quality assurance (QA). \emph{Community}: Standardization and impartiality (governance, transparency, building trust, long term maintainability). \emph{Knowledge}: research and meta-research, academic publications.}
    \label{fig:layers}
\end{figure}

@@

\begin{figure}
    \centering
        %\includegraphics[width=0.75\linewidth]{../original_submission/figures/bens_formalization.pdf}
    \includegraphics[width=0.75\linewidth]{./figures/fig3_bens_formalization_arrows.pdf} % alternative
    \caption[Benchmark formalization.]{Benchmark formalization. \textbf{A}) Running a benchmark requires running a workflow (execution phase) to generate results to be critically evaluated (analysis phase). \textbf{B}) the execution phase consists on a mapping of methods to specific input files to generate output files, with optional steps to harmonize data and parametrize runs.}
    \label{fig:formalization}
\end{figure}

### 2025-04-23 1_first_review/tracked_changes.tex
before generating the bbl

\begin{figure}
    \centering
    \includegraphics[width=\linewidth]{review_cycle_1/figures/fig1_fig_benchmarking_aligned.pdf}
    {\color{red} \caption[Run types, properties, capabilities and features.]{Solo and community benchmarks, workflows, and benchmarking-system-enabled (\textit{bs}) benchmarks all differ in run properties and their consequences in terms of capabilities and features, including whether: inputs, truths, hidden truths, metrics or run results are available for readers; code is free software, software environments are managed, and execution automated; there is a clear track of (human) contributors; results are run in comparable hardware; an entrypoint to results or result interpretation are provided. These run-intrinsic properties have consequences in terms of features (i.e., transparency, forkability, availability for metaanalysis) and technical capabilities (i.e. software versioning or ability to run locally without leaving traces - in \textit{incognito} mode). Green, blue and red ticks depict available, perhaps available and unavailable features, respectively. \label{fig:definitions}}
    }
\end{figure}

@@

\begin{figure}
    \centering
    \includegraphics[width=0.8\linewidth]{../original_submission/figures/fig2_bens_layers.pdf}
    \caption[Benchmarking conceptual layers and their challenges.]{Benchmarking conceptual layers and their challenges. The process of benchmarking is multilayered, with layer-specific factors belonging to different categories (colored boxes) and challenges (in red). \emph{Hardware}: infrastructure and its costs. \emph{Data}: dataset archival, openness, interoperability, and selection.  \emph{Software}: method implementations, reproducibility, workflow execution, continuous integration and delivery (CI/CD), versioning, quality assurance (QA). \emph{Community}: Standardization and impartiality (governance, transparency, building trust, long term maintainability). \emph{Knowledge}: research and meta-research, academic publications.}
    \label{fig:layers}
\end{figure}

>>>>

\begin{figure}
    \centering
    \includegraphics[width=\linewidth]{./figures/fig1_fig_benchmarking_aligned.pdf}
    {\color{red} \caption[Run types, properties, capabilities and features.]{Solo and community benchmarks, workflows, and benchmarking-system-enabled (\textit{bs}) benchmarks all differ in run properties and their consequences in terms of capabilities and features, including whether: inputs, truths, hidden truths, metrics or run results are available for readers; code is free software, software environments are managed, and execution automated; there is a clear track of (human) contributors; results are run in comparable hardware; an entrypoint to results or result interpretation are provided. These run-intrinsic properties have consequences in terms of features (i.e., transparency, forkability, availability for metaanalysis) and technical capabilities (i.e. software versioning or ability to run locally without leaving traces - in \textit{incognito} mode). Green, blue and red ticks depict available, perhaps available and unavailable features, respectively. \label{fig:definitions}}
    }
\end{figure}

@@

\begin{figure}
    \centering
    \includegraphics[width=0.8\linewidth]{./figures/fig2_bens_layers.pdf}
    \caption[Benchmarking conceptual layers and their challenges.]{Benchmarking conceptual layers and their challenges. The process of benchmarking is multilayered, with layer-specific factors belonging to different categories (colored boxes) and challenges (in red). \emph{Hardware}: infrastructure and its costs. \emph{Data}: dataset archival, openness, interoperability, and selection.  \emph{Software}: method implementations, reproducibility, workflow execution, continuous integration and delivery (CI/CD), versioning, quality assurance (QA). \emph{Community}: Standardization and impartiality (governance, transparency, building trust, long term maintainability). \emph{Knowledge}: research and meta-research, academic publications.}
    \label{fig:layers}
\end{figure}
